---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Systematic measurement changes: A simulation study

<!-- badges: start -->
<!-- badges: end -->

This repository contains the code to reproduce the simulation study in the paper  
*"Assessing systematic changes in measurements across a data collection: A 
simulation study"*.

The study compares seven statistical methods for quantifying systematic changes in measurements across different settings:  
- Autoregressive integrated moving average (ARIMA)  
- Fused lasso signal approximator (FLSA)  
- Generalized additive model (GAM)  
- Locally weighted scatterplot smoothing (LOWESS)  
- Moving average (MA)  
- Pruned exact linear time (PELT)  
- Piecewise regression (PR)

The simulation results can be interactively explored via an accompanying [Shiny application](https://sys-measurement-changes.bips.eu/). The source code for the app is available in a separate [GitHub repository](https://github.com/bips-hb/shiny-sys-measurement-changes).

## 📁 Repository structure

- `functions/`: Contains function definitions for [01_simulation_study.R](01_simulation_study.R):
  - [algorithms.R](functions/algorithms.R): Statistical methods
  - [problems.R](functions/problems.R): Simulation settings  
- `renv/`: Contains project-specific environment configuration files.
- `results/`: Contains all outputs of the simulation study, including figures, tables, and supplementary data. Raw results are stored in [res.rds](results/res.rds).  
- `01_simulation_study.R`: Runs the simulation study using the `batchtools` R package. Results are saved in [res.rds](results/res.rds)
- `02_visualisations.R`: Generates [figures](results/figures) and [tables](results/tables) for the main manuscript and [supplementary_data_1.pdf](results/supplementary_data_1.pdf).
- `03_supplementary_data_2.R`: Creates the supplementary Excel file [supplementary_data_2.xlsx](results/supplementary_data_2.xlsx) containing bias and mean squared error for different settings.
- `04_supplementary_data_3.Rmd`: R Markdown to create [supplementary_data_3.pdf](results/supplementary_data_3.pdf) with the output from the [Shiny application](https://sys-measurement-changes.bips.eu/).
- `renv.lock`: Records package versions.

## 🚀 Reproducing the results

To reproduce the simulation study, run [01_simulation_study.R](01_simulation_study.R). This executes the simulations and saves the results to `results/res.rds`.

To recreate the figures and tables for the main manuscript and [supplementary_data_1.pdf](results/supplementary_data_1.pdf),
run [02_visualisations.R](02_visualisations.R). This saves the figures and supplementary figures under `results/figures` and the tables under `results/tables`. 

To recreate the supplementary files  [supplementary_data_2.xlsx](results/supplementary_data_2.xlsx) and [supplementary_data_3.pdf](results/supplementary_data_3.pdf), run [03_supplementary_data_2.R](03_supplementary_data_2.R) and [04_supplementary_data_3.Rmd](04_supplementary_data_3.Rmd), respectively. 

> Note: It is **not necessary** to run the simulation study to create the figures, tables, or supplementary data, as the results are already stored in `results/res.rds`.

## 📚 R environment

This project uses the `renv` R package to ensure reproducible R environments. The exact package version used for the simulation study are recorded in [renv.lock](renv.lock). To restore the environment and install all required packages:

### 1. Clone the repository

Clone the repository (or download the ZIP) to save [renv.lock](renv.lock) locally.

### 2. Open the project

Open the project in R (e.g. in RStudio). Make sure [renv.lock](renv.lock) is in the project root directory.

### 3. Install renv

Install `renv` if not already installed.

```{r, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
install.packages("renv")
```

### 4. Restore project library

Restore the project library using [renv.lock](renv.lock)

```{r, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
renv::restore()
```

